---
title: "ENVS 193DS Infographic"
date: "May 27, 2026" 
author: "Danielle Sur" 
format: pdf 
editor: source
---

# Set up

```{r}
#| message: false

# loaded in packages
library(tidyverse)
library(janitor)

# read in data and saved as object called `rxntime`
rxntime <- read_csv("rxn-time.csv")
```


# Part 3. Code your visualizations

## a. Set up your repository

```{r}
# created a new object from halibut data frame
halibut_summary <- halibut |> 
  # grouped by treatment type
  group_by(ca_region) |> 
  # calculated values for each region
  summarize(
    # summary for parametric test
    mean = mean(landings_lb), # mean
    n = length(landings_lb), # number of observations (n)
    sd = sd(landings_lb), # standard deviation
    # summary for non-parametric test
    count = n(), # count
    median = median(landings_lb), # median
    IQR = IQR(landings_lb) # interquartile range
  )

# display of data frame
halibut_summary
```

### Strip plot for Welch's two-sample t-test

```{r}
# base layer: ggplot
ggplot(data = halibut_summary,
       # x-axis: region
       # y-axis: mean
       mapping = aes(x = ca_region,
                     y = mean,
                     color = ca_region)) +
  # added bars for standard deviation
  geom_errorbar(mapping = aes(ymin = mean - sd,
                              ymax = mean + sd,
                              width = 0.1)) +
  # relabeled axes titles
  labs(x = "Region of California",
       y = "Landings (lb)") +
  # custom theme
  theme_minimal() +
  # removed legend
  theme(legend.position = "none") +
  # raw individual data points (jitter with horizontal break)
  # x-axis: region
  # y-axis: landing weight
  geom_jitter(data = halibut,
            mapping = aes(x = ca_region,
                          y = landings_lb,
                          color = ca_region),
            width = 0.2,
            height = 0,
            size = 1.75,
            alpha = 0.5,
            shape = 18) +
  # added points for the means
  geom_point(size = 3,
             color = "red") +
  # custom colors
  scale_color_manual(values = c("Northern" = "hotpink",
                                "Southern" = "orange")) +
  # removed grid lines
  theme(panel.grid = element_blank(),
        axis.line = element_line(color = "black"))
```

### Box plot for Mann-Whitney U test

```{r}
# base layer: ggplot
ggplot(data = halibut,
       # x-axis: region
       # y-axis: landing weight
       mapping = aes(x = ca_region,
                     y = landings_lb,
                     fill = ca_region)) +
  # first layer creating boxplots for each region
  geom_boxplot() +
  # relabeled axes titles
  labs(x = "Region of California",
       y = "Landings (lb)") +
  # raw individual data points (jitter with horizontal break)
  # x-axis: region
  # y-axis: landing weight
   geom_jitter(data = halibut,
            mapping = aes(x = ca_region,
                          y = landings_lb,
                          color = ca_region),
            width = 0.3,
            height = 0,
            size = 1.5,
            alpha = 0.5,
            shape = 21) +
  # custom theme
  theme_minimal() +
  # removed legend
  theme(legend.position = "none") +
  # custom colors: raw data points
  scale_color_manual(values =
                       c("Northern" = "hotpink",
                         "Southern" = "orange")) +
  # custom colors: boxplot fill
  scale_fill_manual(values = c("Northern" = "hotpink",
                                "Southern" = "orange")) +
  # removed grid lines
  theme(panel.grid = element_blank(),
        axis.line = element_line(color = "black"))
```

## b. Arrange visuazliations using the design platform of your choosing.

ok

# Part 4. Write-up

**General information about design and visualization**

What patterns are you highlighting in each visualization? Why?

Outside of the data visualization, what aesthetic choices (e.g. color, font, arrangement) did you make and why?

In what way do your choices contribute to a compelling or interesting narrative?

**Sources and process**

For each visualization, describe what examples inspired that visualization (cite the author)

Describe your coding process, for example (not the only options for this prompt): did you have to clean up/summarize the data before you started? Which geoms did you start with? What geoms did you end up using, and why?

Describe the tools you used: other people’s code? Google/StackOverflow? ChatGPT (with examples of prompts)?